AN EFFICIENT LOCALIZED MODEL ORDER REDUCTION FRAMEWORK FOR THE SHAPE OPTIMIZATION OF ADDITIVELY MANUFACTURED LATTICE STRUCTURES\break ECCOMAS CONGRESS 2024

### 2024-08-30 paper/paper.tex
add conclusion

\usepackage{tikz}
\usepackage{standalone}
\usepackage{booktabs}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.18}
\usetikzlibrary{%
	calc,
	patterns,
	positioning,
	decorations.pathreplacing,%
	decorations.pathmorphing,%
	arrows.meta,%
}
\usepackage{bamcolors}

@@

\subsection{Global solution procedure} % (fold)
\label{sub:Online efficiency}
With the global GFEM space, each coarse grid cell (\ie{} unit cell of the lattice) can be viewed as a high order FE.
For instance, assuming a coarse grid partition as shown in~\cref{fig:oversampling_domain}, the coarse grid cell is a quadrilateral cell with special ansatz functions $\bm{\psi}_i$.
The components of the FE matrix of the $i-$th coarse grid cell are given by
\begin{align}
    \begin{split}
        a^i_{kl} &= \int_{\varOmega^{\p}_i} \left(\pdv{\bm{\psi}_k}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\vdot\!\vdot \mathbb{C} \vdot\!\vdot
    \left(\pdv{\bm{\psi}_l}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\det(\bm{F}_{\m})\dd{\bm{x}^{\p}}\,,\\
                 &= ({V}^\top)_{km} {K}_{mn}(\mu){V}_{nl}\,.
    \end{split}
    \label{eq:element_matrix}
\end{align}
Here, $\bm{V}$ is the matrix that holds the FE vectors of the GFEM functions $\bm\psi_i$ as column vectors and $\bm{K}(\mu)$ is the stiffness matrix of the subdomain $\varOmega^{\p}_i$ (\ie{} unit cell).
Since $\bm{K}$ is dependent on the parameter $\m$, integration over the full fine scale mesh of the unit cell is required whenever the model is evaluated for a new parameter value which will drastically reduce the computational efficiency of our method.

@@

\subsection{Graded concrete slab} % (fold)
\label{sub:Graded concrete slab}
The numerical example investigated here is motivated by \textit{graded concrete} in which voids are created by inserting hollow bodies into the concrete formwork with the goal of mass reduction, see~\gradientenbetonurl.
The problem is simplified and we consider a two dimensional beam under plane stress conditions as shown in~\cref{fig:beam_sketch}.
The coarse grid partition of half of the beam body $\varOmega^{\p}_{\gl}=[0, 10a]\times[0, a]$ is given by $10$ quadrilateral cells of size $a$, where $a$ is the unit length of the unit cell.
Each unit cell, \ie{} subdomain $\varOmega^{\p}_i$, is taken as a square of size $a$ with a circular void with radius $R$ and is partitioned in \added{384} isoparametric quadrilateral FEs of polynomial degree two, see~\cref{fig:oversampling_domain}\,(b).
With regard to the geometrical parametrization, we consider the variation of the radius $R$ in each subdomain.
For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathcal{P}_{\gl}$ has parameter components $\m_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
For the parent domain we set $\bar{\m}_i=0.2a, i=1, \ldots, 10$ as the reference value.
The point load is distributed as a constant traction over the length of one unit cell (on the far left).

@@

For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
The error in the displacements is measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
\begin{equation}
    e_u(\m) := \frac{\norm{\bm{u}_{\fom} - \bm{u}_{\rom}}}{\norm{\bm{u}_{\fom}}}\,.
    \label{eq:error_u}
\end{equation}
Apart from the displacement field as the primary unknown, we are interested in the first principal component of the \textsc{Cauchy} stress as this will be used to define a stress constraint for the optimization problem.
The error in the stress is defined as
\begin{equation}
    e_{\sigma}(\m) = \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\vert \mathrm{max}(\vb{s}_{\fom})\vert}\,.
    \label{eq:error_sigma}
\end{equation}
Here, $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.
The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
With regard to the error in the displacement, the average value is much smaller than the maximum over the validation set.
In both cases, for local basis size ca. $> 75$ the error decreases only very slowly with increasing number of basis functions.
This is due to the rather moderate size of the training set, as \eg{} $N_{\train}=200$ is used for a local parameter space $\mathcal{P}^k=[R_{-}, R_{+}]^4$ for $k=3, \ldots, 9$.
However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.

@@

\subsection{Shape optimization} % (fold)
\label{sub:Shape optimization}

\begin{itemize}
    \item problem definiot
    \item output computation (difference with ROM)
    \item result for both models
\end{itemize}

@@

\section{CONCLUSIONS} % (fold)
\label{sec:conclusions}

@@

% \begin{figure}[!htb]
% \minipage{0.49\textwidth}
%   \includegraphics[]{./figures/beam/fig_loc_svals_left.pdf}
%   \caption{Singular valus left}\label{fig:loc_svals_left}
% \endminipage\hfill
% \minipage{0.49\textwidth}
%   \includegraphics[]{./figures/beam/fig_loc_svals_right.pdf}
%   \caption{Singular valus right}\label{fig:loc_svals_right}
% \endminipage\hfill\\%
% \begin{center}
% \minipage{0.49\textwidth}
%   \includegraphics[]{./figures/beam/fig_loc_svals_inner.pdf}
%   \caption{Singular valus inner}\label{fig:loc_svals_inner}
% \endminipage
% \end{center}
% \end{figure}

% \begin{figure}[!htb]
% \minipage{0.49\textwidth}
%   \includegraphics{./figures/beam/fig_proj_error_left_hapod.pdf}
%   \caption{Projection error hapod}\label{fig:proj_error_left_hapod}
% \endminipage\hfill
% \minipage{0.49\textwidth}
%   \includegraphics{./figures/beam/fig_proj_error_right_hapod.pdf}
%   \caption{Projection error hapod}\label{fig:proj_error_right_hapod}
% \endminipage\hfill\\
% \begin{center}
% \minipage{0.49\textwidth}%
%   \includegraphics{./figures/beam/fig_proj_error_inner_hapod.pdf}
%   \caption{Projection error hapod}\label{fig:proj_error_inner_hapod}
% \endminipage
% \end{center}
% \end{figure}

% \begin{table}
%     \centering
%     \caption{HAPOD inner}\label{tab:hapod_inner}
%     \input{./src/beam/tables/hapod_inner.tex}
% \end{table}

% \begin{figure}[!htb]
% \minipage{0.49\textwidth}
%   \includegraphics{./figures/beam/fig_proj_error_left_heuristic.pdf}
%   \caption{Projection error }\label{fig:proj_error_left_heuristic}
% \endminipage\hfill
% \minipage{0.49\textwidth}
%   \includegraphics{./figures/beam/fig_proj_error_right_heuristic.pdf}
%   \caption{Projection error }\label{fig:proj_error_right_heuristic}
% \endminipage\hfill\\
% \begin{center}
% \minipage{0.49\textwidth}%
%   \includegraphics{./figures/beam/fig_proj_error_inner_heuristic.pdf}
%   \caption{Projection error }\label{fig:proj_error_inner_heuristic}
% \endminipage
% \end{center}
% \end{figure}

% \begin{table}
%     \centering
%     \caption{Heuristic rrf inner}\label{tab:heuristic_inner}
%     \input{./src/beam/tables/heuristic_inner.tex}
% \end{table}

% \begin{figure}[!htb]
%     \centering
%     \includegraphics{./figures/beam/fig_loc_rom_error.pdf}
%     \caption{Local ROM error relative to FOM.}\label{fig:loc_rom_error}
% \end{figure}

% \begin{table}
%     \caption{Result of the optimization with FOM and ROM.}\label{tab:minimization_data}
%     \centering
%     \input{./src/beam/tables/minimization_data.tex}
% \end{table}

% \begin{table}
%     \caption{Comparison of reduced optimal solution $\mu_N^{\ast}$ and true optimal solution $\mu^{\ast}$.}\label{tab:minimization_comparison}
%     \centering
%     \input{./src/beam/tables/minimization_comparison.tex}
% \end{table}

>>>>

\usepackage{tikz}
\usepackage{standalone}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.18}
\usetikzlibrary{%
	calc,
	patterns,
	positioning,
	decorations.pathreplacing,%
	decorations.pathmorphing,%
	arrows.meta,%
}
\usepackage{bamcolors}

@@

\subsection{Global solution procedure} % (fold)
\label{sub:Online efficiency}
With the global GFEM space, each coarse grid cell (\ie{} unit cell of the lattice) can be viewed as a high order FE.
For instance, assuming a coarse grid partition as shown in~\cref{fig:oversampling_domain}, the coarse grid cell is a quadrilateral cell with special ansatz functions $\bm{\psi}_i$.
The components of the FE matrix of the $i-$th coarse grid cell are given by
\begin{align}
    \begin{split}
        a^i_{kl} &= \int_{\varOmega^{\p}_i} \left(\pdv{\bm{\psi}_k}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\vdot\!\vdot \mathbb{C} \vdot\!\vdot
    \left(\pdv{\bm{\psi}_l}{\bm{x}^{\p}}\vdot\bm{F}^{-1}_{\m}\right)\det(\bm{F}_{\m})\dd{\bm{x}^{\p}}\,,\\
                 &= ({V}^\top)_{km} {K}_{mn}(\m){V}_{nl}\,.
    \end{split}
    \label{eq:element_matrix}
\end{align}
Here, $\bm{V}$ is the matrix that holds the FE vectors of the GFEM functions $\bm\psi_i$ as column vectors and $\bm{K}(\m)$ is the stiffness matrix of the subdomain $\varOmega^{\p}_i$ (\ie{} unit cell).
Since $\bm{K}$ is dependent on the parameter $\m$, integration over the full fine scale mesh of the unit cell is required whenever the model is evaluated for a new parameter value which will drastically reduce the computational efficiency of our method.

@@

\subsection{Graded concrete slab} % (fold)
\label{sub:Graded concrete slab}
The numerical example investigated here is motivated by \textit{graded concrete} in which voids are created by inserting hollow bodies into the concrete formwork with the goal of mass reduction, see~\gradientenbetonurl.
The problem is simplified and we consider a two dimensional beam under plane stress conditions as shown in~\cref{fig:beam_sketch}.
The coarse grid partition of half of the beam body $\varOmega^{\p}_{\gl}=[0, 10a]\times[0, a]$ is given by $10$ quadrilateral cells of size $a$, where $a$ is the unit length of the unit cell.
Each unit cell, \ie{} subdomain $\varOmega^{\p}_i$, is taken as a square of size $a$ with a circular void with radius $R$ and is partitioned in 384 isoparametric quadrilateral FEs of polynomial degree two, see~\cref{fig:oversampling_domain}\,(b).
With regard to the geometrical parametrization, we consider the variation of the radius $R$ in each subdomain.
For the global problem this results in a parameter space $\mathcal{P}_{\gl}=[R_{-}, R_{+}]^{10}$, such that a parameter value $\m\in\mathcal{P}_{\gl}$ has parameter components $\mu_i\in[R_{-},R_{+}], i=1, \ldots, 10$, with ranges chosen as $R_{-}=0.1a, R_{+}=0.3a$.
For the parent domain we set $\bar{\mu}_i=0.2a, i=1, \ldots, 10$ as the reference value.
The point load is distributed as a constant traction over the length of one unit cell (on the far left).

@@

For the validation of the ROM, the relative error with respect to the FOM is computed over a parameter set of size $200$.
The error in the displacements is measured in the energy norm (over the domain $\varOmega^{\p}_{\gl}$) and defined as
\begin{equation}
    e_u(\m) := \frac{\norm{\bm{u}_{\fom} - \bm{u}_{\rom}}}{\norm{\bm{u}_{\fom}}}\,.
    \label{eq:error_u}
\end{equation}
Apart from the displacement field as the primary unknown, we are interested in the first principal component of the \textsc{Cauchy} stress as this will be used to define a stress constraint for the optimization problem.
The error in the stress is defined as
\begin{equation}
    e_{\sigma}(\m) = \frac{\vert\vb{s}_{\fom} - \vb{s}_{\rom}\vert}{\vert \mathrm{max}(\vb{s}_{\fom})\vert}\,.
    \label{eq:error_sigma}
\end{equation}
Here, $\vb{s}\in\mathbb{R}^{N_g}$ denotes the vector holding the first principal stress values at the $N_g$ integration points of the mesh.
For ease of notation, the parameter-dependence is omitted for $\bm{u}$ and $\vb{s}$.
The minimum, average and maximum of $e_u$ and $e_{\sigma}$ over the validation set are shown in~\cref{fig:relerr}.
With regard to the error in the displacement, the average value is much smaller than the maximum over the validation set.
In both cases, for local basis size ca. $> 75$ the error decreases only very slowly with increasing number of basis functions.
This is due to the rather moderate size of the training set, as \eg{} $N_{\train}=200$ is used for a local parameter space $\mathcal{P}^k=[R_{-}, R_{+}]^4$ for $k=3, \ldots, 9$.
However, the error in the stress is less sensitive to variations in the parameter $\m$ and a pointwise maximum relative error below $1\%$ is acceptable for the current application.
% \comment[id=pd]{Actually, this is probably only because the ROM stress is computed on the full mesh but with reduced $u$ as input. Reduced stress computation?}

@@

\subsection{Shape optimization} % (fold)
\label{sub:Shape optimization}
The formulation of the shape optimization problem seeks to minimize an objective function, $J(\m, \bm{u}(\m))$, with constraints, $g_q(\m, \bm{u}(\m)) \le 0$, used to limit the first principal stress at each integration point, $q$, with $q=1, \ldots, N_g$.
\begin{align}
    \begin{split}
        \min_{\m\in\mathcal{P}_{\gl}} &J(\m, \bm{u}(\m))\,,\\
        \mathrm{s.t.}\quad & g_q(\m, \bm{u}(\m))\le 0, \quad q=1, \ldots, N_q\,,\\
        \mathrm{with}\quad &\bm{K}(\m)\bm{u}(\m)=\bm{f}\,.
    \end{split}
    \label{eq:shape_opt}
\end{align}
The objective function is defined as a weighted sum of mass $m(\m)$ and compliance $C(\bm{u}(\m))$.
\begin{equation}
    J(\m, \bm{u}(\m))= (1-\omega)\frac{m(\m)}{m(\bar\m)}+\omega \frac{C(\bm{u}(\m))}{C(\bm{u}(\bar\m))}\,,\quad 0\le \omega\le 1\,.
    \label{eq:output}
\end{equation}
The compliance is included, since we do not actively constrain the second principal component and, therefore, designs leading to high compressive stresses would not be detected.
The reference values $m(\bar\m)$ and $C(\bm{u}(\bar\m))$ are thus used to normalize.
The definition of the stress constraint follows a risk factor approach~\cite{Pastore2019Topology} and is written as
\begin{equation}
    g_q(\m, \bm{u}(\m)) = \frac{\vb{s}_q}{\sigma_{+}} - 1\,,\quad q=1,\ldots,N_q\,,
    \label{eq:constraint}
\end{equation}
with the components of the vector of first principal stress $\vb{s}_q$ and an upper bound $\sigma_{+}$ on the admissible tensile stress.
Note that $\vb{s}$ depends on $\bm{u}$ and thus also $\m$ although not explicitly written.

The shape optimization problem~\cref{eq:shape_opt} is then solved using the FOM and the ROM.
Note that the computation of the ROM output $J_n(\m, \bm{u}_n(\m))$, with reduced displacement solution $\bm{u}_n(\m)$, is not dependent on the dimension of the FOM.
The results are obtained using sequential least squares programming~\cite{kraft1988software} (SLSQP) and finite difference approximation of the gradients, and are given in~\cref{tab:opt_result}.
% The relative error in the output $J$ is $\approx\,0.4\%$, while the relative error in the mass $m$ is $\approx\,0.5\%$.
The absolute error in the optimum is $\sqrt{\sum_{i=1}^{10}(\mu_i^{\ast}-\mu_{n,i}^{\ast})^2} \approx 2.44\cdot 10^{-2}$, with the optimum, $\m^{\ast}$, found by the FOM and the optimum, $\m_n^{\ast}$, found by the ROM, respectively.
While the ROM is able to converge to a solution close to the FOM solution, over three times more iterations and function evaluations are necessary.

\begin{table}
    \caption{Result of the optimization. For the ROM a local basis size of $100$ is used. $\m^{\ast}$ is used as a placeholder for the optimum found by the respective model.}\label{tab:opt_result}
    \begin{center}
        \begin{tabularx}{\textwidth}{XXXXX}
            \toprule
            Model & $J(\m^{\ast})$ & $m(\m^{\ast})$ & Iterations & Function calls\\\midrule
            FOM   &  0.941         &  7.855  & 4          & 45\\
            ROM   &  0.946         &  7.897  & 13         & 155\\
            \bottomrule
        \end{tabularx}
    \end{center}
\end{table}

@@

\section{CONCLUSIONS} % (fold)
\label{sec:conclusions}
In this contribution a hyper-reduced localized model order reduction framework for the shape optimization of lattice structures is presented.
The data compression strategy, adopted from~\cite{Taddei2018Localization}, is shown to work well with range approximation by random sampling.
Furthermore, the ROM solution of the considered shape optimization problem is in good agreement with the FOM solution and this is regarded as a first proof of concept.
A thorough comparison of the computational efficiency including runtime measurements is planned for future work, as this is out of the scope of the current investigation.
Other directions of future research include the exploration of offline training stategies in which the variation in the parameter value $\m$ and the boundary data $\bm{g}$ are addressed simultaneously, instead of sequentially as this may lead to redundant data in the snapshot set $\bm{S}$.

@@
